Renal IV

### 2017-08-20 RESP2.rmd
sed -i 's/=>/$\rightarrow$/g' *

CO2 retainers: chronic increased PaCO2 => decreased sensitivity in the central chemoceptors => hypoxia becomes main breathing drive => DON'T GIVE O2

@@

Loss of lung tissue => decreases lung recoil => increased intrapleural pressure => early airway closure, decreased FVC, FEV1. It INCREASES PIF due to decreased recoil.

>>>>

CO2 retainers: chronic increased PaCO2 $\rightarrow$ decreased sensitivity in the central chemoceptors $\rightarrow$ hypoxia becomes main breathing drive $\rightarrow$ DON'T GIVE O2

@@

Loss of lung tissue $\rightarrow$ decreases lung recoil $\rightarrow$ increased intrapleural pressure $\rightarrow$ early airway closure, decreased FVC, FEV1. It INCREASES PIF due to decreased recoil.

### 2017-08-20 msk2.rmd
Updated yaml header to make it a4. Stuff american letter paper
sed -i 's/=>/$\rightarrow$/g' *

```
title: "MSK II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: pdf_document
```

@@

-  Matrix composed of 35% osteoid (mainly type I collagen) and 65% mineral component (mainly hydroxyapatite).
-  2 types of matrix = {woven, lamellar}
-  Cells = {osteoblasts, osteoclasts, osteocytes}
-  Development:
    -  Long bones from endochondral ossification
    -  Flat bones from intramembranous ossification
-  Remodelling:
    -  RANKL: stimulates RANK => increased osteoclast differentiation (stimulated by PTH)
    -  OPG: inhibits RANKL RANK signalling => decreased osteoclast differentiation

@@

-  Dysostoses: defects in migration and condensation of mesenchyme
-  Dysplasia: global disorganisation of bone/cartilage
    -  Achondroplasia: most common AD disorder from g.o.f *FGFR3* mutation => retarded cartilage growth
    -  Thanatophoric dysplasia: lethal dwarfism also g.o.f *FGFR3*
    -  Osteogenesis imperfecta: mutation in *type I collagen* => dominant negative lof mutation => too little bone => fragility
    -  Osteopetrosis: reduced bone resorption resulting from impaired formation or fn of osteoclasts. AR mutations in *CA2* or *TCIRG1*

@@

1.  Initial osteolytic stage
2.  Mixed osteoclastic/osteoblastic stage => predominance of osteoblastic activity
3.  Burned out quiescent osteosclerotic phase

@@

-  Rupture of vessles results in haematoma surrounding area of bone injury
-  Clotted blood provides fibrin mesh, sealing off fracture site, and providing scaffold for inflamm cells/fibroblasts/angiogenesis.
-  Degradnulated platelets and inflamm cells release PDGF, TGF-$\beta$, FGF to activate osteoprogenitor cells.
-  End of first week: mass of uncalcified tissue called *soft callus* provides anchorage between ends of fractured bone.
-  After two weeks: soft callus => bony callus due to deposition of woven bone, hyaline/fibro-cartilage, and endochondral ossification.

@@

According to criteria she'd have 6 points => definite RA

Natural Hx: progressive deformities of jt, 10% increased risk of CVS mortality within 10 years of Dx (inflamm => vasculitis/athero)

@@

-  Inflamm of tendons (such as calcaneal tendon)
-  Oedema
-  Ossification
-  Fusion of spine => bamboo spine

@@

-  Hyperuricaemia: underexcretion (90%) or overproduction (10%)
-  Excess urate => crystalisation => inflamm => phago of crystals => inflamm: joints get affected
-  Can clear up in a couple of days

@@

-  Acute
    -  NSAIDs
    -  Prednisone
    -  Colchicine for acute gout: interrupts cycle of mono-Na crystal deposition. Stops the inflammatory reaction to the crystals. 
-  Chronic
    -  Lower urate to $<0.36\text{mmol/L}$
    -  Lifestyle modifications: low purine diet, low sugar (fructose => hyperuricaemia), stay hydrated
    -  Allopurinol: inhibits xanthine oxidase => inhibits urate production. Not really given during acute flare up because it *may* exacerbates.

>>>>

---
title: "MSK II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
  pdf_document
documentclass: article
classoption: a4paper
---

@@

-  Matrix composed of 35% osteoid (mainly type I collagen) and 65% mineral component (mainly hydroxyapatite).
-  2 types of matrix = {woven, lamellar}
-  Cells = {osteoblasts, osteoclasts, osteocytes}
-  Development:
    -  Long bones from endochondral ossification
    -  Flat bones from intramembranous ossification
-  Remodelling:
    -  RANKL: stimulates RANK $\rightarrow$ increased osteoclast differentiation (stimulated by PTH)
    -  OPG: inhibits RANKL RANK signalling $\rightarrow$ decreased osteoclast differentiation

@@

-  Dysostoses: defects in migration and condensation of mesenchyme
-  Dysplasia: global disorganisation of bone/cartilage
    -  Achondroplasia: most common AD disorder from g.o.f *FGFR3* mutation $\rightarrow$ retarded cartilage growth
    -  Thanatophoric dysplasia: lethal dwarfism also g.o.f *FGFR3*
    -  Osteogenesis imperfecta: mutation in *type I collagen* $\rightarrow$ dominant negative lof mutation $\rightarrow$ too little bone $\rightarrow$ fragility
    -  Osteopetrosis: reduced bone resorption resulting from impaired formation or fn of osteoclasts. AR mutations in *CA2* or *TCIRG1*

@@

1.  Initial osteolytic stage
2.  Mixed osteoclastic/osteoblastic stage $\rightarrow$ predominance of osteoblastic activity
3.  Burned out quiescent osteosclerotic phase

@@

-  Rupture of vessles results in haematoma surrounding area of bone injury
-  Clotted blood provides fibrin mesh, sealing off fracture site, and providing scaffold for inflamm cells/fibroblasts/angiogenesis.
-  Degradnulated platelets and inflamm cells release PDGF, TGF-$\beta$, FGF to activate osteoprogenitor cells.
-  End of first week: mass of uncalcified tissue called *soft callus* provides anchorage between ends of fractured bone.
-  After two weeks: soft callus $\rightarrow$ bony callus due to deposition of woven bone, hyaline/fibro-cartilage, and endochondral ossification.

@@

According to criteria she'd have 6 points $\rightarrow$ definite RA

Natural Hx: progressive deformities of jt, 10% increased risk of CVS mortality within 10 years of Dx (inflamm $\rightarrow$ vasculitis/athero)

@@

-  Inflamm of tendons (such as calcaneal tendon)
-  Oedema
-  Ossification
-  Fusion of spine $\rightarrow$ bamboo spine

@@

-  Hyperuricaemia: underexcretion (90%) or overproduction (10%)
-  Excess urate $\rightarrow$ crystalisation $\rightarrow$ inflamm $\rightarrow$ phago of crystals $\rightarrow$ inflamm: joints get affected
-  Can clear up in a couple of days

@@

-  Acute
    -  NSAIDs
    -  Prednisone
    -  Colchicine for acute gout: interrupts cycle of mono-Na crystal deposition. Stops the inflammatory reaction to the crystals. 
-  Chronic
    -  Lower urate to $<0.36\text{mmol/L}$
    -  Lifestyle modifications: low purine diet, low sugar (fructose $\rightarrow$ hyperuricaemia), stay hydrated
    -  Allopurinol: inhibits xanthine oxidase $\rightarrow$ inhibits urate production. Not really given during acute flare up because it *may* exacerbates.

### 2017-08-20 neuro2.rmd
sed -i 's/=>/$\rightarrow$/g' *

-  UMN => increased external sphincter tone
-  LMN => bowel cannot physically move the stool along so it gets stuck

@@

Pressure sores: pressure on area of tissue => lack of perfusion => anaerobic metabolism => tissue ischaemia. Spinal injuries: can't feel it.

@@

High doses on muscle: fasciculations => inability to repolarise cell membranes => weakness + paralysis => ventilatory failure =>

@@

-  MSK: strain on erector spinae, infn of vertebrae, poor hip posture => functional scoliosis, spondylothesis (fracture), spondylitis, spinal stenosis, vertebral fracture, scoliosis, disc prolapse, osteomyelitis, epidural abscesses (IV drug users), OA
-  CVS: aortic aneurysm
-  GIT: pancreatitis, gall bladder infn, colon Ca, kidney stones, ischaemia intestines
-  Pregnancy stuff
-  Metabolic: cushings, acromegaly, paget disease
-  Miscellaneous: herpes zosta, sarcoidosis

>>>>

-  UMN $\rightarrow$ increased external sphincter tone
-  LMN $\rightarrow$ bowel cannot physically move the stool along so it gets stuck

@@

Pressure sores: pressure on area of tissue $\rightarrow$ lack of perfusion $\rightarrow$ anaerobic metabolism $\rightarrow$ tissue ischaemia. Spinal injuries: can't feel it.

@@

High doses on muscle: fasciculations $\rightarrow$ inability to repolarise cell membranes $\rightarrow$ weakness + paralysis $\rightarrow$ ventilatory failure $\rightarrow$

@@

-  MSK: strain on erector spinae, infn of vertebrae, poor hip posture $\rightarrow$ functional scoliosis, spondylothesis (fracture), spondylitis, spinal stenosis, vertebral fracture, scoliosis, disc prolapse, osteomyelitis, epidural abscesses (IV drug users), OA
-  CVS: aortic aneurysm
-  GIT: pancreatitis, gall bladder infn, colon Ca, kidney stones, ischaemia intestines
-  Pregnancy stuff
-  Metabolic: cushings, acromegaly, paget disease
-  Miscellaneous: herpes zosta, sarcoidosis

### 2017-08-20 neuro3.rmd
sed -i 's/=>/$\rightarrow$/g' *

-  Trauma: haematoma
-  Vascular: aneurysm, haemorrhage => $\Uparrow$ neck stiffness, hypertensive emergency
-  Inflamm: meningitis (may not be infn), encephalitis, MS
-  Malnutrition
-  Migraine: first presentation often gets CT'd
-  Hydrocephalus: accumulation of CSF
-  Tumour => bleeding => mass effect
-  Cyst
-  Iatrogenic: medicines can cause headaches

@@

> 29M falling off bike, head first into ground. Pain back of head and neck, neuro rapidly degenerates as he gets to the hospital going from responsive => unresponsive

>>>>

-  Trauma: haematoma
-  Vascular: aneurysm, haemorrhage $\rightarrow$ $\Uparrow$ neck stiffness, hypertensive emergency
-  Inflamm: meningitis (may not be infn), encephalitis, MS
-  Malnutrition
-  Migraine: first presentation often gets CT'd
-  Hydrocephalus: accumulation of CSF
-  Tumour $\rightarrow$ bleeding $\rightarrow$ mass effect
-  Cyst
-  Iatrogenic: medicines can cause headaches

@@

> 29M falling off bike, head first into ground. Pain back of head and neck, neuro rapidly degenerates as he gets to the hospital going from responsive $\rightarrow$ unresponsive

### 2017-08-20 renal4.rmd
sed -i 's/=>/$\rightarrow$/g' *

Neurogenic bladder: problems with the nervous system causing urinary retention / incontinence. Have to empty this to prevent stasis => infn.

@@

Mechanism for dizziness when straining to void: increasing abdominal pressure => stimulation of vagal nerve => micturition syncope (emptying large amounts of V from bladder).

@@

Paediatrics is like veterinary science => have to ask random q's since child can't explain things themselves. \textbf{Always ask more and examine more in children}.

@@

-  Ca
-  Septic shock
-  Meningococcal disease (Neiserria meningitis (gram neg coccus); meningitis): neck stiffness/rigidity, fever, koenig's sign, Brudzinski's sign (more for adults), rash (late sign; caused by septic emboli getting into tissue and causing microinfarctions => non-blanching rash which separates it from dermal rashes), decreased level of consciousness
-  Measles: C's = {koplic spots (white spots in mouth), conjuctivitis, coryza}, fever

>>>>

Neurogenic bladder: problems with the nervous system causing urinary retention / incontinence. Have to empty this to prevent stasis $\rightarrow$ infn.

@@

Mechanism for dizziness when straining to void: increasing abdominal pressure $\rightarrow$ stimulation of vagal nerve $\rightarrow$ micturition syncope (emptying large amounts of V from bladder).

@@

Paediatrics is like veterinary science $\rightarrow$ have to ask random q's since child can't explain things themselves. \textbf{Always ask more and examine more in children}.

@@

-  Ca
-  Septic shock
-  Meningococcal disease (Neiserria meningitis (gram neg coccus); meningitis): neck stiffness/rigidity, fever, koenig's sign, Brudzinski's sign (more for adults), rash (late sign; caused by septic emboli getting into tissue and causing microinfarctions $\rightarrow$ non-blanching rash which separates it from dermal rashes), decreased level of consciousness
-  Measles: C's = {koplic spots (white spots in mouth), conjuctivitis, coryza}, fever
